---
title: "Florida COVID-19 Hospitalisations"
author: "Gabriel J. Odom, Roy Williams, Zoran Bursac, and Mary Jo Trepka"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r packages}
# Load packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(kableExtra)
```

# Hospitalisation Data
```{r hosptitalised}
# Import Data
library(readxl)
data_dir <- "../../data/"
dataFiles_char <- list.files(path = data_dir, pattern = "ESS_.*hrs")
# NOTE: on and after 10 June, they removed the "s" from "hrs". I've added it to
#   the file names manually (as I have to save the files manually from my email
#   anyway, I can add the "s" to the file name).

read_excel_safely <- safely(read_excel)

floridaHospitalised_ls <-
	map(
		.x = dataFiles_char,
		.f = ~{

			# The tab name changed in June (month 06)
			sheetName_char <- case_when(
	    	str_detect(
	    		.x, pattern = "Hospitals_04"
	    	) ~ "County_Bed_Availability_Report_",
	    	str_detect(
	    		.x, pattern = "Hospitals_05"
	    	) ~ "County_Bed_Availability_Report_",
	    	TRUE ~ "County_Bed_Avail_Report_for_Par"
	    )
			
			# The sheet header changed on June 20th.
			skip0_lgl <- str_detect(.x, "06.20.2020")
			
	    df_ls <- read_excel_safely(
	      path = paste0(data_dir, .x),
	      sheet = sheetName_char,
        skip = ifelse(test = skip0_lgl, yes = 0, no = 1)
      )

	    if(!is.null(df_ls$error)){
        NULL
	    } else {
	    	
	    	df_ls$result %>%
    	    filter(County == "Grand Total") %>%
    	    select(contains("COVID")) %>%
    	    mutate(
    	    	Date_char = str_remove(
    	    		.x,
    	    		pattern = "ESS_BedAvailabilityWithAddlinfo_Hospitals_"
    	    	)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_sub(Date_char, end = 10)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_replace_all(
    	    		Date_char,
    	    		pattern = "\\.",
    	    		replacement = "-"
    	    	)
    	    ) %>%
    	    mutate(Date = as.POSIXct(Date_char, format = "%m-%d-%Y")) %>%
    	    select(-Date_char) %>%
    	    select(Date, everything())
	    	
	    }

		}
	)
```

```{r}
# Wrangle Data
# We had a data wrangling error, so I wrapped read_excel() in a safely() and
#   added an if() statement.
# floridaHospitalised_ls %>%
# 	map_lgl(is.null) %>%
# 	which()
# 30 83 84 85 86 87
# dataFiles_char[30]
# "ESS_BedAvailabilityWithAddlinfo_Hospitals_04.23.2020_1045hrs.xlsx"
# I inspected this file manually and found the sheet's name was abbreviated
floridaHospitalised_ls[[30]] <-
	read_excel(
    path = paste0(data_dir, dataFiles_char[30]),
    sheet = "County",
    skip = 1
  ) %>%
  filter(County == "Grand Total") %>%
  select(contains("COVID")) %>%
  mutate(Date_char = "2020-04-23") %>%
  mutate(Date = as.POSIXct(Date_char, format = "%Y-%m-%d")) %>%
  select(-Date_char) %>%
  select(Date, everything())

# dataFiles_char[83]
# After June 1, they changed the sheet name.
# floridaHospitalised_ls[[83]]


floridaHospitalised_df <-
	floridaHospitalised_ls %>%
	bind_rows() %>%
	mutate(Date = as_date(Date)) %>% 
	mutate(Hospitalised = `COVID IN ICU` + `COVID NON ICU`) %>%
  rename(
  	ICU = `COVID IN ICU`,
  	Ventilated = `COVID ON VENT`,
  	AdmitPrevDay = `COVID + Admits Day Before`,
  	DischPrevDay = `COVID + Disch Day Before`
  ) %>%
	mutate(DeltaAdmit = AdmitPrevDay - DischPrevDay) %>%
  select(
  	Date, Hospitalised, ICU, Ventilated, AdmitPrevDay, DischPrevDay, DeltaAdmit
  )
```

```{r}
# Calculate Dates and Save
startEndESS_date <- 
	floridaHospitalised_df %>%
	slice(1, n()) %>%
	pull(Date) 

# write_csv(
# 	floridaHospitalised_df,
# 	path = paste0(
# 		"../data/ESS_Florida_summary_",
# 		format(startEndESS_date[2], "%Y%m%d"),
# 		".csv"
# 	)
# )
```


## Full Data
```{r}
# Display Data
knitr::kable(floridaHospitalised_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

The data above is COVID-19 hospital census data in Miami-Dade county for all dates with available data since April 2. These columns are: the number of patients hospitalized with COVID-19 (**Hospitalised**), the number of patients in intensive care with COVID-19 (**ICU**), the number of patients in intensive care with COVID-19 who are also on ventilators (**Ventilated**), the number of patients positive for COVID-19 that were admitted the previous day (**AdmitPrevDay**), the number of patients recovered from COVID-19 that were discharged the previous day (**DischPrevDay**), and the net admission of patients positive for COVID-19 the previous day (**DeltaAdmit**; a negative value means more recovered patients were discharged than sick patients admitted). *There is a data error for the count ventilated on the 4th of June; we do not know the true value, but we were told verbally by local Department of Health officials that the data point should be approximately 55 for Miami-Dade. We removed this point from the data before estimating the models shown below.*

```{r pivot_hosp_data}
floridaHospitalised_df[
	floridaHospitalised_df$Date == "2020-06-04",
	"Ventilated"
] <- NA_real_
# There is an error in the original data for June 4th. Rather than deleting the
#   whole row of the raw data, we are skipping this ventilation value

floridaCOVID_df <-
	floridaHospitalised_df %>%
	pivot_longer(
		Hospitalised:DeltaAdmit,
		names_to = "Type",
		values_to = "Count"
	)
```


## Hospital COVID-19 Census
We now show a census plot of hospitalizations, patients in ICU, and those in ICU on ventilators (shown on a log scale):
```{r graph_hosp_counts, warning = FALSE}
ggplot(
	data = floridaCOVID_df %>%
		filter(Type != "DischPrevDay") %>%
		filter(Type != "AdmitPrevDay") %>%
		filter(Type != "DeltaAdmit") 
) +

	theme_bw() +
	theme(legend.position = "bottom") + 
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"Ventilated" = "#ff0000",
			"ICU" = "#ff7400",
			"Hospitalised" = "#ffc100",
			"AdmitPrevDay" = "black"
		)
	) +
	scale_y_log10() +
	labs(
		title = "Overall Florida Hospital COVID-19 Census",
		subtitle = paste(
			"State of Florida;",
			format(startEndESS_date[1], "%d %B"),
			"to",
			format(startEndESS_date[2], "%d %B"),
			"2020"
		),
		y = "Counts (Log10 Scale)"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)
```

Overall, since mid-April, the number of COVID-19 positive patients in Miami-Dade County hospitals has remained constant at around 700 people in hospitals. The COVID-19 ICU census has decreased from its reported peak of 285 on April 17th to around 100 as of June 6th. Overall, the number of COVID-19 patients on ventilators has decreased significantly from its highest point of 192 ventilated on April 22nd to around 60 on June 6th. However, **the observed decreasing trends for ICU census and ventilation census have reversed over the past week**. Note: data is plotted on a log scale.


*******************************************************************************


# Conclusion

Public understanding of science is essential in times of public health crisis. We offer a few closing remarks.

- **The number of positive tests obtained tends to increase as the overall number of tests performed increases.** Simply stated, more tests result in more opportunities for positive results. As an analogy, overall more students will pass a class if more students are enrolled in the class. If we are interested in the difficulty of a course, the percentage of students that pass the class is a better gauge of difficulty than the number of students who pass the class. This is because some classes are large, and some classes are small.
- **Based on the increasing number of COVID-19 tests performed in Miami-Dade County, the proportion of positive tests obtained is the best metric to access COVID-19 epidemic progression.** Overall, the proportion of positive tests obtained has decreased from a high of 20% in early April, to the current rate of between 4-5%. Over the last two weeks, however, the proportion has slightly increased. However, this percentage is still (currently) holding near the 5% threshold recommended by many experts for reopening. Similarly, the number of COVID-19 positive individuals in the hospital, ICU, and on ventilators at the same time has substantially decreased since mid April. However, over the last 2 weeks, the numbers are no longer decreasing, but have seemed to flatten out. From our estimation, hospitals in Miami-Dade County still have sufficient resources to accommodate potential increases.
- **These metrics will change over time based on human behavior.** Because of this, these metrics will be closely monitored to assess the effect of reopening policies.


*******************************************************************************


# Data Sources
Data sources are as follows:

- **COVID-19 Test Results**: The number of positive and negative COVID-19 test results come from the Florida Department of Health: https://floridahealthcovid19.gov/.
- **COVID-19 Hospital Census**: The hospitalization data is disseminated by Florida’s Agency for Health Care Administration: https://ahca.myflorida.com/.


