---
title: "S. Florida COVID-19 Trajectory"
author: "Gabriel J. Odom, PhD, ThD"
date: "5/29/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Analyse Daily COVID-19 Trajectory  

```{r packages, message=FALSE}
library(tidyverse)
library(lubridate)
```


## Testing Data
```{r read_data, message=FALSE}
sflCases_df <- read_csv(
	file = "../../data/FLDH_COVID19_cases_20200529.csv"
)
```

```{r}
knitr::kable(sflCases_df)
```

```{r fix_dates, message=FALSE}
sflCases2_df <- 
	sflCases_df %>% 
	mutate(Date = as.POSIXct(strptime(Date, format = "%d-%b"))) %>% 
	mutate(Date = as_date(Date)) 
```

We see the following pattern of positive and negative tests over the past month (shown on a log scale):
```{r cases_over_time, message=FALSE}
ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date) + 
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	scale_y_log10() +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = "2 April to 28 May, 2020",
		caption = "Green represents a positive test result while blue is negative.",
		y = "Counts (Log10 Scale)"
	) +
	
	geom_point(aes(y = Positive), colour = "green") +
	stat_smooth(aes(y = Positive), method = "loess", colour = "green") +
  geom_point(aes(y = Negative), colour = "blue") +
	stat_smooth(aes(y = Negative), method = "loess", colour = "blue")
```

### Data Leading up to Making the Decision to Open
The proportion of positive tests over the two weeks before we decided to open the economy is (this is the data used to make the desicion to partially open the economy in Miami-Dade):
```{r prop_over_time_before_open, message=FALSE}
# Linear Trend Slope
slope1 <- sflCases2_df %>% 
	filter(Date > "2020-04-19") %>% 
	filter(Date < "2020-05-05") %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Plot
ggplot(
	data = sflCases2_df %>% 
		filter(Date > "2020-04-19") %>% 
		filter(Date < "2020-05-05")
) +
	
	theme_bw() +
	aes(x = Date) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = paste(
			"20 April to 4 May, 2020; Linear Change =",
			round(slope1, 3)
		),
		caption = "Solid line represents the LOESS Predictor; Dashed line for Linear.",
		y = "Proportion of Positive Cases"
	) +
	
	geom_point(aes(y = PropPositive)) +
	stat_smooth(aes(y = PropPositive), method = "loess", colour = "black") +
	stat_smooth(
		aes(y = PropPositive),
		method = "lm",
		colour = "black",
		se = FALSE,
		linetype = "dashed"
	)
```

### Data Between Publicising the Decision to Open and Opening
The proportion of positive tests over the two weeks before we actually opened the economy is (this is what happened between when we decided to open the economy and declared it open):
```{r prop_over_time_interim, message=FALSE}
# Linear Trend Slope
slope2 <- sflCases2_df %>% 
	filter(Date > "2020-04-26") %>% 
	filter(Date < "2020-05-12") %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Plot
ggplot(
	data = sflCases2_df %>% 
		filter(Date > "2020-04-26") %>% 
		filter(Date < "2020-05-12")
) +
	theme_bw() +
	aes(x = Date) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = paste(
			"27 April to 11 May, 2020; Linear Change =",
			round(slope2, 3)
		),
		caption = "Solid line represents the LOESS Predictor; Dashed line for Linear.",
		y = "Proportion of Positive Cases"
	) +
	geom_point(aes(y = PropPositive)) +
	stat_smooth(aes(y = PropPositive), method = "loess", colour = "black") +
	stat_smooth(
		aes(y = PropPositive),
		method = "lm",
		colour = "black",
		se = FALSE,
		linetype = "dashed"
	)
```

### Data During the Open Economy at First
The proportion of positive tests in the last two weeks before we actually opened the economy is (this is what happened when we declared the economy was open):
```{r prop_over_time_after_open, message=FALSE}
# Linear Trend Slope
slope3 <- sflCases2_df %>% 
		filter(Date > "2020-05-03") %>% 
		filter(Date < "2020-05-19") %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Plot
ggplot(
	data = sflCases2_df %>% 
		filter(Date > "2020-05-03") %>% 
		filter(Date < "2020-05-19")
) +
	theme_bw() +
	aes(x = Date) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = paste(
			"4 May to 18 May, 2020; Linear Change =",
			round(slope3, 3)
		),
		caption = "Solid line represents the LOESS Predictor; Dashed line for Linear.",
		y = "Proportion of Positive Cases"
	) +
	geom_point(aes(y = PropPositive)) +
	stat_smooth(aes(y = PropPositive), method = "loess", colour = "black") +
	stat_smooth(
		aes(y = PropPositive),
		method = "lm",
		colour = "black",
		se = FALSE,
		linetype = "dashed"
	)
```

### Recent Data During the Open Economy
The proportion of positive tests in the last two weeks is:
```{r prop_over_time_now, message=FALSE}
# Linear Trend Slope
slope4 <- sflCases2_df %>% 
	filter(Date > "2020-05-16") %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Plot
ggplot(
	data = sflCases2_df %>% 
		filter(Date > "2020-05-16") 
) +
	theme_bw() +
	aes(x = Date) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = paste(
			"16 May 2020 onward; Linear Change =",
			round(slope4, 3)
		),
		caption = "Solid line represents the LOESS Predictor; Dashed line for Linear.",
		y = "Proportion of Positive Cases"
	) +
	geom_point(aes(y = PropPositive)) +
	stat_smooth(aes(y = PropPositive), method = "loess", colour = "black") +
	stat_smooth(
		aes(y = PropPositive),
		method = "lm",
		colour = "black",
		se = FALSE,
		linetype = "dashed"
	)
```

### Full Data
These are the proportions of positive tests for the full data.
```{r prop_over_time_full, message=FALSE}
ggplot(data = sflCases2_df) +
	
	theme_bw() +
	aes(x = Date) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 in Miami Dade County",
		subtitle = "2 April to 31 May, 2020",
		caption = "Solid line represents the LOESS Predictor; Dashed line for Linear.",
		y = "Proportion of Positive Cases"
	) +
	
	geom_point(aes(y = PropPositive)) +
	stat_smooth(aes(y = PropPositive), method = "loess", colour = "black") +
	stat_smooth(
		aes(y = PropPositive),
		method = "lm",
		colour = "black",
		se = FALSE,
		linetype = "dashed"
	)
```



## Hospitalisation Data
<!-- We haven't received new ESS data for two weeks. I'm cutting this section. -->
```{r hosptitalised}
library(readxl)
data_dir <- "../../data/"
dataFiles_char <- list.files(path = data_dir, pattern = "ESS_.*hrs")

read_excel_safely <- safely(read_excel)

miamidadeHospitalised_ls <-
	map(
		.x = dataFiles_char,
		.f = ~{

	    df_ls <- read_excel_safely(
	      path = paste0(data_dir, .x),
        sheet = "County_Bed_Availability_Report_",
        skip = 1
      )

	    if(is.null(df_ls$error)){

	    	df_ls$result %>%
    	    filter(County == "MIAMI-DADE") %>%
    	    select(contains("COVID")) %>%
    	    mutate(
    	    	Date_char = str_remove(
    	    		.x,
    	    		pattern = "ESS_BedAvailabilityWithAddlinfo_Hospitals_"
    	    	)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_sub(Date_char, end = 10)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_replace_all(
    	    		Date_char,
    	    		pattern = "\\.",
    	    		replacement = "-"
    	    	)
    	    ) %>%
    	    mutate(Date = as.POSIXct(Date_char, format = "%m-%d-%Y")) %>%
    	    select(-Date_char) %>%
    	    select(Date, everything())

	    } else {
	    	NULL
	    }

		}
	)

# We had a data wrangling error, so I wrapped read_excel() in a safely() and
#   added an if() statement.
# miamidadeHospitalised_ls %>%
# 	map_lgl(is.null) %>%
# 	which()
# 30
# dataFiles_char[30]
# "ESS_BedAvailabilityWithAddlinfo_Hospitals_04.23.2020_1045hrs.xlsx"
# I inspected this file manually and found the sheet's name was abbreviated
miamidadeHospitalised_ls[[30]] <-
	read_excel(
    path = paste0(data_dir, dataFiles_char[30]),
    sheet = "County",
    skip = 1
  ) %>%
  filter(County == "MIAMI-DADE") %>%
  select(contains("COVID")) %>%
  mutate(Date_char = "2020-04-23") %>%
  mutate(Date = as.POSIXct(Date_char, format = "%Y-%m-%d")) %>%
  select(-Date_char) %>%
  select(Date, everything())


miamidadeHospitalised_df <-
	miamidadeHospitalised_ls %>%
	bind_rows() %>%
	mutate(Hospitalised = `COVID IN ICU` + `COVID NON ICU`) %>%
  	rename(
  		ICU = `COVID IN ICU`,
  		Ventilated = `COVID ON VENT`,
  		AdmitPrevDay = `COVID + Admits Day Before`,
  		DischPrevDay = `COVID + Disch Day Before`
  	) %>%
	mutate(DeltaAdmit = AdmitPrevDay - DischPrevDay) %>%
  select(
  	Date, Hospitalised, ICU, Ventilated, AdmitPrevDay, DischPrevDay, DeltaAdmit
  )
```

```{r}
knitr::kable(miamidadeHospitalised_df)
```


```{r pivot_hosp_data}
mdCOVID_df <-
	miamidadeHospitalised_df %>%
	pivot_longer(
		Hospitalised:DeltaAdmit,
		names_to = "Type",
		values_to = "Count"
	)
```

### Hosptial COVID-19 Census
Census plot of hospitalisations, patients in ICU, and those in ICU on ventilators.
```{r graph_hosp_counts, warning = FALSE, message=FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type != "DischPrevDay") %>%
		filter(Type != "AdmitPrevDay") %>%
		filter(Type != "DeltaAdmit")
) +

	theme_bw() +
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"Ventilated" = "#ff0000",
			"ICU" = "#ff7400",
			"Hospitalised" = "#ffc100",
			"AdmitPrevDay" = "black"
		)
	) +
	scale_y_log10() +
	labs(
		title = "Counts of Patients Hospitalised with COVID-19 in Miami Dade County",
		subtitle = "2 April to 31 May, 2020",
		y = "Counts (Log10 Scale)"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)
```

### Net Hospital Admissions
#### Net New Admissions
```{r graph_hosp_net_admit, warning = FALSE, message=FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type == "DeltaAdmit")
) +

	theme_bw() +
	aes(x = Date, y = Count) +
	labs(
		title = "COVID-19: New Hospital Admissions minus Discharges",
		subtitle = "Miami Dade County: 2 April to 31 May, 2020",
		y = "Net Admits"
	) +

	geom_point(size = 2) +
	geom_hline(yintercept = 0, colour = "red") +
	stat_smooth(se = FALSE)

```

#### Admissions vs Dicharges
```{r graph_hosp_admit_disch, warning = FALSE, message=FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type %in% c("DischPrevDay", "AdmitPrevDay"))
) +

	theme_bw() +
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"AdmitPrevDay" = "#a61414",
			"DischPrevDay" = "#0C479D"
		)
	) +
	labs(
		title = "Counts of New Hospital Admissions and Discharges with COVID-19",
		subtitle = "Miami Dade County: 2 April to 31 May, 2020",
		y = "Counts"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)

```

