---
title: "South Florida COVID-19 Trajectory"
author: "Gabriel J. Odom, Roy Williams, Zoran Bursac, and Mary Jo Trepka"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Analyse Daily COVID-19 Trajectory  

```{r packages, message=FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)
library(kableExtra)
```

**INTRODUCTION HERE**


*******************************************************************************


# Testing Data
```{r read_data, message=FALSE}
sflCases_df <- read_csv(
	file = "../../data/FLDH_COVID19_cases_20200607.csv"
)
```

```{r}
knitr::kable(sflCases_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

**EXPLAIN DATA HERE**

```{r fix_dates, message=FALSE}
sflCases2_df <- 
	sflCases_df %>% 
	mutate(Date = as.POSIXct(strptime(Date, format = "%d-%b"))) %>% 
	mutate(Date = as_date(Date)) 

startEnd_date <- 
	sflCases2_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")
```

## Positive Counts to Total Counts
We see the following pattern of positive and negative tests over the past two months (total counts shown on a log scale):
```{r cases_over_time, message=FALSE}
nTotal_gg <- 
	ggplot(
		data = sflCases2_df %>%
			mutate(Total = Positive + Negative)
	) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Total) + 
	scale_y_log10(
		breaks = c(2000, 5000, 10000),
		labels = c("2k", "5k", "10k")
	) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Counts of Positive and Total COVID-19 Tests in Miami Dade",
		subtitle = paste(
			startEnd_date[1], "to", startEnd_date[2], "2020"
		),
		y = "Counts (Log10 Scale)"
	) +
	
	stat_smooth(method = "loess", colour = "black") +
  geom_point(colour = "black")


nPositive_gg <- 
	ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Positive) + 
	# scale_y_log10() +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		caption = "Green represents a positive test result while black is total."
	) + 
	
	stat_smooth(method = "loess", colour = "green") +
	geom_point(colour = "green") 


plot_grid(
	nTotal_gg, nPositive_gg,
	nrow = 2,
	labels = "AUTO"
) 
```

**EXPLAIN FIGURES HERE**


## Proportion of Positive Cases within Past Two Weeks
The proportion of positive tests for the past two weeks is:
```{r recent_data, message=FALSE}
recent_df <- 
	sflCases2_df %>% 
	top_n(15, Date)

# Date Range
recentStartEnd_date <- 
	recent_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")

# Linear Trend Slope
recentSlope_num <- recent_df %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)
```

```{r recent_proportion, message=FALSE}
ggplot(data = recent_df) +
	
	theme_bw() +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 Tests in Miami Dade County",
		subtitle = paste(
			recentStartEnd_date[1], "to", recentStartEnd_date[2],
			"2020; Linear Change =",
			round(recentSlope_num, 3)
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Proportion of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "black") +
	stat_smooth(method = "lm", colour = "black", se = FALSE, linetype = "dashed")
```

**EXPLAIN FIGURE HERE**


## Proportion of Positive Cases Overall
These are the proportions of positive tests for the full data.
```{r prop_over_time_full, message=FALSE}
ggplot(data = sflCases2_df) +
	
	theme_bw() +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive COVID-19 Tests in Miami Dade County",
		subtitle = paste(
			startEnd_date[1], "to", startEnd_date[2], "2020"
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Proportion of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "black") +
	stat_smooth(method = "lm", colour = "black", se = FALSE, linetype = "dashed")
```

**EXPLAIN FIGURE HERE**


*******************************************************************************


# Hospitalisation Data
<!-- Should we cut this section? -->
```{r hosptitalised}
library(readxl)
data_dir <- "../../data/"
dataFiles_char <- list.files(path = data_dir, pattern = "ESS_.*hrs")
# NOTE: on and after 10 June, they removed the "s" from "hrs". I've added it to
#   the file names manually (as I have to save the files manually from my email
#   anyway, I can add the "s" to the file name).

read_excel_safely <- safely(read_excel)

miamidadeHospitalised_ls <-
	map(
		.x = dataFiles_char,
		.f = ~{

	    df_ls <- read_excel_safely(
	      path = paste0(data_dir, .x),
        sheet = ifelse(
        	# The tab name changed in June (06)
        	test = str_detect(.x, pattern = "Hospitals_06"),
        	yes = "County_Bed_Avail_Report_for_Par", 
        	no = "County_Bed_Availability_Report_"
        ),
        skip = 1
      )

	    if(!is.null(df_ls$error)){
        NULL
	    } else {
	    	
	    	df_ls$result %>%
    	    filter(County == "MIAMI-DADE") %>%
    	    select(contains("COVID")) %>%
    	    mutate(
    	    	Date_char = str_remove(
    	    		.x,
    	    		pattern = "ESS_BedAvailabilityWithAddlinfo_Hospitals_"
    	    	)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_sub(Date_char, end = 10)
    	    ) %>%
    	    mutate(
    	    	Date_char = str_replace_all(
    	    		Date_char,
    	    		pattern = "\\.",
    	    		replacement = "-"
    	    	)
    	    ) %>%
    	    mutate(Date = as.POSIXct(Date_char, format = "%m-%d-%Y")) %>%
    	    select(-Date_char) %>%
    	    select(Date, everything())
	    	
	    }

		}
	)
```

```{r}
# We had a data wrangling error, so I wrapped read_excel() in a safely() and
#   added an if() statement.
# miamidadeHospitalised_ls %>%
# 	map_lgl(is.null) %>%
# 	which()
# 30 83 84 85 86 87
# dataFiles_char[30]
# "ESS_BedAvailabilityWithAddlinfo_Hospitals_04.23.2020_1045hrs.xlsx"
# I inspected this file manually and found the sheet's name was abbreviated
miamidadeHospitalised_ls[[30]] <-
	read_excel(
    path = paste0(data_dir, dataFiles_char[30]),
    sheet = "County",
    skip = 1
  ) %>%
  filter(County == "MIAMI-DADE") %>%
  select(contains("COVID")) %>%
  mutate(Date_char = "2020-04-23") %>%
  mutate(Date = as.POSIXct(Date_char, format = "%Y-%m-%d")) %>%
  select(-Date_char) %>%
  select(Date, everything())

# dataFiles_char[83]
# After June 1, they changed the sheet name.
# miamidadeHospitalised_ls[[83]]


miamidadeHospitalised_df <-
	miamidadeHospitalised_ls %>%
	bind_rows() %>%
	mutate(Date = as_date(Date)) %>% 
	mutate(Hospitalised = `COVID IN ICU` + `COVID NON ICU`) %>%
  rename(
  	ICU = `COVID IN ICU`,
  	Ventilated = `COVID ON VENT`,
  	AdmitPrevDay = `COVID + Admits Day Before`,
  	DischPrevDay = `COVID + Disch Day Before`
  ) %>%
	mutate(DeltaAdmit = AdmitPrevDay - DischPrevDay) %>%
  select(
  	Date, Hospitalised, ICU, Ventilated, AdmitPrevDay, DischPrevDay, DeltaAdmit
  )
```


```{r}
startEndESS_date <- 
	miamidadeHospitalised_df %>%
	slice(1, n()) %>%
	pull(Date) 

# write_csv(
# 	miamidadeHospitalised_df,
# 	path = paste0(
# 		"../data/ESS_MiamiDade_summary_",
# 		format(startEndESS_date[2], "%Y%m%d"),
# 		".csv"
# 	)
# )

knitr::kable(miamidadeHospitalised_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

**EXPLAIN DATA HERE**  
There is a data error for the count ventilated on the 4th of June.


```{r pivot_hosp_data}
mdCOVID_df <-
	miamidadeHospitalised_df %>%
	pivot_longer(
		Hospitalised:DeltaAdmit,
		names_to = "Type",
		values_to = "Count"
	)
```

## Hosptial COVID-19 Census
Census plot of hospitalisations, patients in ICU, and those in ICU on ventilators.
```{r graph_hosp_counts, warning = FALSE, message=FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type != "DischPrevDay") %>%
		filter(Type != "AdmitPrevDay") %>%
		filter(Type != "DeltaAdmit") %>% 
		filter(Date != "2020-06-04")
	# There is an error in the original data for June 4th. Rather than changing
	#   the raw data, we are skipping this date
) +

	theme_bw() +
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"Ventilated" = "#ff0000",
			"ICU" = "#ff7400",
			"Hospitalised" = "#ffc100",
			"AdmitPrevDay" = "black"
		)
	) +
	scale_y_log10() +
	labs(
		title = "Counts of Patients Hospitalised with COVID-19 in Miami Dade County",
		subtitle = paste(
			format(startEndESS_date[1], "%d %B"),
			"to",
			format(startEndESS_date[2], "%d %B"),
			"2020"
		),
		y = "Counts (Log10 Scale)"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)
```

**EXPLAIN FIGURE HERE**


*******************************************************************************


# Conclusion

**CONCLUSION / DISCUSSION HERE**