---
title: "Miami-Dade County COVID-19 Trajectory"
author: "Roy Williams, Gabriel J. Odom, Zoran Bursac, and Mary Jo Trepka"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


# Introduction

Public understanding of science is critical in times of crisis. The following analysis was conducted to examine the trajectory of COVID-19 in Miami-Dade County. For phased reopening, the White House has set guidelines for local municipalities in their “Opening Up American Again” report. In the report, it is highly suggested that:

1. Local areas show a downward trajectory in the number of positive cases as a percent of total tests over a two-week period, and
2. Local hospitals have capacity to handle all patients without crisis care.

The analysis begins by looking at the number of new positive cases each day and the number of total tests performed each day. The figures are presented together because **as there is more testing, there are likely to be more positive test results**. Therefore, the proportion of positive cases is also calculated. Given either a flat or increasing number of tests performed, **the proportion of positive cases is the key metric to be used when looking at epidemic trajectory**. It is recommended that this proportion <ins>not exceed</ins> 5%.

Finally, to ensure hospitals can meet demand without crisis care, the COVID-19 positive hospitalization census, COVID-19 positive ICU census, and COVID-19 positive ventilation census is plotted over time. Census refers to the number of COVID-19 positive patients in the hospital, ICU, and on ventilators, respectively, on a given day. **Note: 5-day moving averages are used to smooth data.**

**Note: a data dump of previously reported results prevented the reporting system from normal processing of lab results on October 9th, causing an inaccurate representation of data for October 9th and October 10th. The most accurate way to interpret this data is to use an a combined average for these days.**

*******************************************************************************


# COVID-19 Test Counts and Proportions
```{r packages}
# Load packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(kableExtra)
library(readxl)
```

```{r rolling_mean_function}
RollUp <- function(.x, .w, .f = mean, ...){
	# Find "Rolling" values of a specified function
	# Inputs:
	#   .x: a vector (usually numeric) over which to "roll" the function .f
	#   .w: the window size
	#   .f: the fuction to "roll" over the values of .x
	#   ...: additional arguments to .f
	# Output: a vector the length of .x with the first (.w - 1) values set to NA
	#   and the remaining values equal to .f evaluated over the previous window.
	# Details: for a five-day moving average, set .w = 5. Then the moving average
	#   of .x at index 5 will be mean(.x[1], .x[2], ..., .x[5]).
	# Examples:
	#   # Rolling mean of first five integers
	#   RollUp(.x = 1:10, .w = 5)
	
	n <- length(.x)
	out <- rep(NA, times = n)
	class(out) <- class(.x)
	
	for(i in .w:n) {
		out[i] <- .f(.x[(i - .w + 1):i])
	}
	
	out
	
}
```


```{r read_data}
# Import data
sflCases_df <- read_csv(
	file = "../../data/cases/FLDH_COVID19_cases_miamidade_20201108.csv"
) %>% 
	mutate(Total = Negative + Positive) %>% 
	mutate(PropPositive = 100 * Positive / Total) %>% 
	# Move proportion to last column
	select(-PropPositive, PropPositive)
```

```{r roll_up_testing_data}
# Rolling Means
sflCasesMA_df <- 
	sflCases_df %>% 
	mutate(Positive = RollUp(.x = Positive, .w = 5, na.rm = TRUE)) %>% 
	mutate(Negative = RollUp(.x = Negative, .w = 5, na.rm = TRUE)) %>% 
	mutate(Total = RollUp(.x = Total, .w = 5, na.rm = TRUE)) %>%
	mutate(PropPositive = 100 * Positive / Total)
```



## Full Data
This is a table of the most recent data we have available. We include data sources at the end of the document.
```{r}
knitr::kable(sflCases_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

The data above show for each day since April 2: the number of positive COVID-19 tests, the number of negative COVID-19 tests, and the proportion of positive COVID-19 tests to total COVID-19 tests.

```{r fix_dates}
sflCases2_df <- 
	sflCasesMA_df %>% 
	mutate(Date = as.POSIXct(strptime(Date, format = "%d-%b"))) %>% 
	mutate(Date = as_date(Date)) 

startEnd_date <- 
	sflCases2_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")
```

```{r recent}
recent_df <- 
	sflCases2_df %>% 
	top_n(15, Date)

# Date Range
recentStartEnd_date <- 
	recent_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")

# Linear Cases Trend Slope
recentCasesSlope_num <-
	recent_df %>% 
	lm(Positive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Adjectives for Cases
adjective_cases <- case_when(
	recentCasesSlope_num > 0.0 ~ "increase",
	recentCasesSlope_num < 0.0 ~ "decrease"
)

adjective_cases_2 <- case_when(
	recentCasesSlope_num > 0.0 ~ "additional",
	recentCasesSlope_num < 0.0 ~ "fewer"
)

# Linear Total Test Trend Slope
recent_TestSlope_num <- 
	recent_df %>% 
	lm(Total ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Slope Adjective for Test:
adjective_tests <- case_when(
	recent_TestSlope_num > 0.0 ~ "an increase",
	recent_TestSlope_num < 0.0 ~ "a decrease"
)

# Slope Adjectives: Additional or Fewer
adjective_tests_2 <- case_when(
	recent_TestSlope_num > 0.0 ~ "additional",
	recent_TestSlope_num < 0.0 ~ "fewer"
)
```


## Total COVID-19 Tests by Day
We see the following pattern for the count of total tests. (shown on a log scale):
```{r tests_over_time}
totalTestSlope_num <- 
	sflCases2_df %>% 
	lm(Total ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

nTotal_gg <- 
	ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Total) + 
	scale_y_log10(
		breaks = c(2000, 5000, 10000),
		labels = c("2k", "5k", "10k")
	) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of COVID-19 Tests by Day; Miami Dade County",
		subtitle = paste(
			startEnd_date[1], "to", startEnd_date[2],
			"2020; Average Two Week Change =",
			round(totalTestSlope_num, 1),
			"Tests Added Daily"
		),
		y = "Counts (Log10 Scale)"
	) +
	
	stat_smooth(method = "loess", colour = "black") +
  geom_point(colour = "black")

nTotal_gg
```

The average number of COVID-19 tests performed each day has increased from about 2,000 in early April to around 12,500 currently. Over the past two weeks there was `r adjective_tests` of `r round(totalTestSlope_num,1)` `r adjective_tests_2` tests performed each day.


## Positive COVID-19 Tests by Day
We see the following pattern for the count of positive tests:
```{r cases_over_time}
nPositive_gg <- 
	ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Positive) + 
	# scale_y_continuous(limits = c(0, 1000)) +
  # scale_y_log10() +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of Positive COVID-19 Tests Results by Day",
		subtitle = paste(
			"Miami-Dade County;",
			startEnd_date[1], "to", startEnd_date[2], "2020"
		),
		y = "Count Positive"
	) + 
	
	stat_smooth(method = "loess", colour = "green") +
	
	geom_point(colour = "green") 

nPositive_gg
```

There was a general decrease in the number of new positive test results from early April until around May 18th, the day Miami-Dade County reopened. Afterwards, this trend flattened out and then increased until late July before it started to decrease. **Over the past two weeks, there has been an average increase of 15.9 additional positive cases each day.** 

Over the past two weeks, there has been an average `r adjective_cases` of `r round(recentCasesSlope_num,0)` `r adjective_cases_2` cases reported each day.

### Positive COVID-19 Tests within the Past Two Weeks

In addition to the overall trend of positive COVID-19 tests shown above, we also show an enhanced view of the last two weeks.
```{r cases_over_two_weeks}
nPositiveRecent_gg <- 
	ggplot(data = recent_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Positive) + 
	# scale_y_continuous(limits = c(0, 1000)) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of Positive Test Results within the Past Two Weeks",
		subtitle = paste(
			"Miami Dade County;",
			recentStartEnd_date[1], "to", recentStartEnd_date[2],
			"2020; Linear Change =",
			round(recentCasesSlope_num, 1)
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Count of Positive Cases"
	) +
	
	stat_smooth(method = "loess", colour = "green") +
	
	geom_point(colour = "green") 


nPositiveRecent_gg
```


## Proportion of Positive Cases within Past Two Weeks
```{r recent_data}
# Linear Trend Slope
recentSlope_num <-
	recent_df %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Slope Adjective:
adjective <- case_when(
	abs(recentSlope_num) < 0.1 ~ "slight",
	abs(recentSlope_num) < 0.5 ~ "",
	abs(recentSlope_num) > 0.5 ~ "significant"
)


# Slope Adjectives: Additional or Fewer

adjective2 <- case_when(
	abs(recentSlope_num) < 0.0 ~ "additional",
	abs(recentSlope_num) > 0.0 ~ "fewer",
)

```

```{r recent_proportion}
ggplot(data = recent_df) +
	
	theme_bw() +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive Test Results within the Past Two Weeks",
		subtitle = paste(
			"Miami Dade County;",
			recentStartEnd_date[1], "to", recentStartEnd_date[2],
			"2020; Linear Change =",
			round(recentSlope_num, 3)
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Proportion (%) of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "black") +
	stat_smooth(method = "lm", colour = "black", se = FALSE, linetype = "dashed")
```

During the 14-day period from `r recentStartEnd_date[1]` to `r recentStartEnd_date[2]`, there was a <b> `r adjective` `r if (recentSlope_num > 0) "increase" else "decrease"` </b> in the proportion of positive COVID-19 cases reported in Miami-Dade County. During this time, there was an average `r adjective_cases` of  `r round(recentSlope_num, 3)`% per day for new positive cases. Given a flat or increasing number of tests performed, the proportion of positive tests obtained is the **key metric** for examining COVID-19 trajectory.


## Proportion of Positive COVID-19 Tests by Day
```{r md_prop_over_time_full}
ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Overall Proportion of Positive Test Results Out of All Tests",
		subtitle = paste(
			"Miami-Dade County;",
			startEnd_date[1], "to", startEnd_date[2], "2020"
		),
		caption = "Solid line: LOESS Predictor; Orange Triangle: Phase 1 Reopening, Red Triangles: Phase 2 and Phase 3 Reopenings",
		y = "Proportion (%) of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "orange") +
  geom_point(
    data = tibble(
      Date = as_date("2020-05-18"),
      y = 0
    ),
    aes(x = Date, y = y),
    pch = 24, size = 2, fill = "orangered"
  )+
geom_point(
		data = tibble(
			Date = as_date("2020-09-21"),
			y = 0
		),
		aes(x = Date, y = y),
		pch = 24, size = 2, fill = "red3"
	)+ 
	geom_point(
		data = tibble(
			Date = as_date("2020-09-25"),
			y = 0
		),
		aes(x = Date, y = y),
		pch = 24, size = 2, fill = "red2"
	)
```

Overall, the proportion of positive test results out of all tests decreased significantly from over 20% in April to 5% on May 18th, the date Miami-Dade County reopened. **After reopening, the proportion of positive tests increased steadily until late July. Subsequently, there was a consistent decline in positivity rate until mid-September. Recently, the change in positivity rate has plateaued.** This number is function of test availability and the population being tested.



*******************************************************************************


# Hospitalization Data
```{r hosptitalised}
# Import data
miamidadeHospitalized_df <- read_csv(
	file = "../../data/ESS_processed/ESS_southFL_summary_20201108.csv"
) %>% 
	filter(County == "MIAMI-DADE")

# Calculate Dates and Save
startEndESS_date <- 
	miamidadeHospitalized_df %>%
	slice(1, n()) %>%
	pull(Date) 


```


## Full Data
```{r}
# Display Data
knitr::kable(miamidadeHospitalized_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

The data above is COVID-19 hospital census data in Miami-Dade County for all dates with available data since April 2. These columns are: the number of patients hospitalized with COVID-19 (**Hospitalized**), the number of patients in intensive care with COVID-19 (**ICU**), the number of patients in intensive care with COVID-19 who are also on ventilators (**Ventilated**), the number of patients positive for COVID-19 that were admitted the previous day (**AdmitPrevDay**), the number of patients recovered from COVID-19 that were discharged the previous day (**DischPrevDay**), and the net admission of patients positive for COVID-19 the previous day (**DeltaAdmit**; a negative value means more recovered patients were discharged than sick patients admitted). Not shown is the number of persons under investigation (PUI).*There is a data error for the count ventilated on the 4th of June. We removed this point from the data before estimating the models shown below.*

```{r pivot_hosp_data}
miamidadeHospitalized_df[
	miamidadeHospitalized_df$Date == "2020-06-04",
	"Ventilated"
] <- NA_real_
# There is an error in the original data for June 4th. Rather than deleting the
#   whole row of the raw data, we are skipping this ventilation value

# Rolling Means
miamidadeHospitalizedMA_df <- 
	miamidadeHospitalized_df %>% 
	mutate(Hospitalized = RollUp(.x = Hospitalized, .w = 5, na.rm = TRUE)) %>% 
	mutate(ICU = RollUp(.x = ICU, .w = 5, na.rm = TRUE)) %>% 
	mutate(Ventilated = RollUp(.x = Ventilated, .w = 5, na.rm = TRUE))

mdCOVID_df <-
	miamidadeHospitalizedMA_df %>%
	pivot_longer(
		Hospitalized:DeltaAdmit,
		names_to = "Type",
		values_to = "Count"
	)
```




## Hospital COVID-19 Census
We now show a census plot of hospitalizations, patients in ICU, and those in ICU on ventilators (shown on a log scale):
```{r graph_hosp_counts, warning = FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type != "DischPrevDay") %>%
		filter(Type != "AdmitPrevDay") %>%
		filter(Type != "DeltaAdmit") 
) +

	theme_bw() +
	theme(legend.position = "bottom") + 
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"Ventilated" = "#ff0000",
			"ICU" = "#ff7400",
			"Hospitalized" = "#ffc100",
			"AdmitPrevDay" = "black"
		)
	) +
	scale_y_log10() +
	labs(
		title = "Overall Miami-Dade County Hospital COVID-19 Census",
		subtitle = paste(
			"Miami Dade County;",
			format(startEndESS_date[1], "%d %B"),
			"to",
			format(startEndESS_date[2], "%d %B"),
			"2020"
		),
		y = "Counts (Log10 Scale)"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)
```

Overall, since mid-April, the number of COVID-19 positive patients in Miami-Dade County hospitals remained constant at around 700 people in hospitals until early June. The COVID-19 ICU census decreased from its reported peak of 285 on April 17th to around 100 as of June 6th. Overall, the number of COVID-19 patients on ventilators decreased from its highest point of 192 ventilated on April 22nd to around 60 as of June 6th. After June 6th, the hospitalizaton burden increased until approximately the beginging of August.

**Recently, on November 6th, there were `r miamidadeHospitalized_df$Hospitalized[nrow(miamidadeHospitalized_df)]` COVID positive patients hospitalized. This was a decrease from a high of 2299 obtained on July 21st. Additionally on November 6th, there were `r miamidadeHospitalized_df$ICU[nrow(miamidadeHospitalized_df)]` COVID positive patients in the ICU. This was a decrease from from a high of 561 obtained on July 26th. Furthermore, on November 6th, there were `r miamidadeHospitalized_df$Ventilated[nrow(miamidadeHospitalized_df)]` COVID positive patients ventilated. This was a decrease from a high of 335 obtained on July 24th. ** Note: data is plotted on a log scale.


*******************************************************************************
# COVID-19 Deaths by Day
```{r}
# Import Data
deathsbyday_df <- read_csv(
	file = "../../data/deaths/FLDH_COVID19_deathsbyday_bycounty_20201108.csv"
) %>% 
	# Deaths are now eight weeks behind
	filter(Date <= "2020-09-15") 

startEndDeaths_date <- 
	deathsbyday_df %>%
	filter(County == "Dade") %>% 
	slice(1, n()) %>%
	pull(Date)
```


### Full Data
```{r md_deaths_data}
whichCounty <- "Dade" # "Palm Beach" // "Broward"

deathsbyday_df %>% 
	filter(County == whichCounty) %>% 
  knitr::kable() %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

```{r}
# Rolling Means
deathsbydayMA_df <- 
	deathsbyday_df %>% 
	filter(County == whichCounty) %>% 
	mutate(Count = RollUp(.x = Count, .w = 5, na.rm = TRUE))
```

```{r}
# Plot Deaths
ggplot(deathsbydayMA_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Count) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "COVID-19 Deaths per Day",
		subtitle = paste(
			"Miami-Dade County;",
			format(startEndDeaths_date[1], "%d %B"),
			"to",
			format(startEndDeaths_date[2], "%d %B"),
			"2020"
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor."
		
	)+
	
	geom_point() +
	stat_smooth(method = "loess", colour = "red") +
	geom_point(
		data = tibble(
			Date = as_date("2020-05-18"),
			y = 0
		),
		aes(x = Date, y = y),
		pch = 24, size = 2, fill = "orange"
	)
```

Overall from March 16th to  approximately May 25th in Miami-Dade County, there was a  decrease in the number of deaths per day. The orange triangle depicts the county reopening date. This decrease in deaths may be attributable to advances in COVID-19 treatments and standard of care. A decrease in median age of those impacted could also partially explain this decrease. **After May 25th, there was an increase in deaths reported per day for Miami-Dade County culminating in a peak during the second week of July. Aftewards, the number of reported deaths per day has started to decrease. The peak number of deaths so far reported was 76 on July 17th.**

**Note: Due to delays in classifying the cause of death and thus of reporting deaths due to COVID-19, the most recent data will undercount the number of deaths. Specifically, 25% of newly reported deaths were from on or before September 13th. This date indicates a point where 75% of our data is considered complete. For this reason, approximately the last 8 weeks were excluded from this analysis of deaths.**

*******************************************************************************

# Conclusion

During the past two weeks, the number of positive COVID-19 tests obtained has **increased** by an average of 15.9 additional positive tests each day. The proportion of positive COVID-19 tests **increased slightly** over the past two weeks. The number of people with COVID-19 who are hospitalize **increased**, while the number in the ICU and on ventilators remained stable. The number of COVID-19 tests performed **increased** for this two-week period. The number of COVID-19 deaths per day decreased.

*******************************************************************************


# Data Sources
Data sources are as follows:

- **COVID-19 Test Results**: The number of positive and negative COVID-19 test results come from the Florida Department of Health: https://urldefense.proofpoint.com/v2/url?u=https-3A__floridahealthcovid19.gov&d=DwIGaQ&c=lhMMI368wojMYNABHh1gQQ&r=ZzssdlRrRS7QXnhuTSa63A&m=OP8LVranCfy0X4-Zh7EQsOWyazaF9sV-b4OHEn301kg&s=oeI6Ns25h9AL46W5e97rQz90vu4ksPvDQt5nYyNuerg&e= .

- **COVID-19 Hospital Census**: The hospitalization data is disseminated by Florida’s Agency for Health Care Administration: https://urldefense.proofpoint.com/v2/url?u=https-3A__ahca.myflorida.com&d=DwIGaQ&c=lhMMI368wojMYNABHh1gQQ&r=ZzssdlRrRS7QXnhuTSa63A&m=OP8LVranCfy0X4-Zh7EQsOWyazaF9sV-b4OHEn301kg&s=YESvxSuxdK9zw0zGUOyiQu1K687UMcfykPUpuyW2QH8&e= .

- **COVID-19 Death Linelist**: This list is created and disseminated by Florida's Department of Health. Deaths are medical examiner confirmed.
https://urldefense.proofpoint.com/v2/url?u=https-3A__floridahealthcovid19.gov&d=DwIGaQ&c=lhMMI368wojMYNABHh1gQQ&r=ZzssdlRrRS7QXnhuTSa63A&m=OP8LVranCfy0X4-Zh7EQsOWyazaF9sV-b4OHEn301kg&s=oeI6Ns25h9AL46W5e97rQz90vu4ksPvDQt5nYyNuerg&e= .


