---
title: "Miami-Dade County COVID-19 Trajectory"
author: "Roy Williams, Gabriel J. Odom, Zoran Bursac, and Mary Jo Trepka"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


# Introduction

Public understanding of science is critical in times of crisis. The following analysis was conducted to examine the trajectory of COVID-19 in Miami-Dade County. For phased reopening, The White House has set guidelines for local municipalities in their “Opening Up American Again” report. In the report, it is highly suggested that:

1. Local areas show a downward trajectory in the number of positive cases as a percent of total tests over a two-week period, and
2. Local hospitals have capacity to handle all patients without crisis care.

The analysis begins by looking at the number of new positive cases each day and the number of total tests performed each day. The figures are presented together because **as there is more testing, there are likely to be more positive test results**. Therefore, the proportion of positive cases is also calculated. Given either a flat or increasing number of tests performed, **the proportion of positive cases is the key metric to be used when looking at epidemic trajectory**. It is recommended that this proportion <ins>not exceed</ins> 5%.

Finally, to ensure hospitals can meet demand without crisis care, the COVID-19 positive hospitalization census, COVID-19 positive ICU census, and COVID-19 positive ventilation census is plotted over time. Census refers to the number of COVID-19 positive patients in the hospital, ICU, and on ventilators, respectively, on a given day.


*******************************************************************************


# COVID-19 Test Counts and Proportions
```{r packages}
# Load packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(kableExtra)
```

```{r read_data}
# Import data
sflCases_df <- read_csv(
	file = "../../data/FLDH_COVID19_cases_20200622.csv"
) %>% 
	mutate(Total = Negative + Positive) %>% 
	mutate(PropPositive = 100 * Positive / (Positive + Negative)) %>% 
	# Move proportion to last column
	select(-PropPositive, PropPositive)
```


## Full Data
This is a table of the most recent data we have available. We include data sources at the end of the document.
```{r}
knitr::kable(sflCases_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

The data above show for each day since April 2: the number of positive COVID-19 tests, the number of negative COVID-19 tests, and the proportion of positive COVID-19 tests to total COVID-19 tests.

```{r fix_dates}
sflCases2_df <- 
	sflCases_df %>% 
	mutate(Date = as.POSIXct(strptime(Date, format = "%d-%b"))) %>% 
	mutate(Date = as_date(Date)) 

startEnd_date <- 
	sflCases2_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")
```


## Total COVID-19 Tests by Day
We see the following pattern for the count of total tests over the past two months (shown on a log scale):
```{r tests_over_time}
totalTestSlope_num <- 
	sflCases2_df %>% 
	lm(Total ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

nTotal_gg <- 
	ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Total) + 
	scale_y_log10(
		breaks = c(2000, 5000, 10000),
		labels = c("2k", "5k", "10k")
	) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of COVID-19 Tests by Day; Miami Dade County",
		subtitle = paste(
			startEnd_date[1], "to", startEnd_date[2],
			"2020; Average Change =",
			round(totalTestSlope_num, 1),
			"Tests Added Daily"
		),
		y = "Counts (Log10 Scale)"
	) +
	
	stat_smooth(method = "loess", colour = "black") +
  geom_point(colour = "black")

nTotal_gg
```

The average number of COVID-19 tests performed each day has doubled from about 2,000 in early April to over 4,500 in mid-June. On average, there is an increase of approximately 34 addition tests performed over the previous day. This data is presented on a log scale.


## Positive COVID-19 Tests by Day
We see the following pattern for the count of positive tests over the past two months:
```{r cases_over_time}
nPositive_gg <- 
	ggplot(data = sflCases2_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Positive) + 
	scale_y_continuous(limits = c(0, 1000)) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of Positive COVID-19 Tests Results by Day",
		subtitle = paste(
			"Miami-Dade County;",
			startEnd_date[1], "to", startEnd_date[2], "2020"
		)
	) + 
	
	stat_smooth(method = "loess", colour = "green") +
	geom_point(colour = "green") 

nPositive_gg
```

There was a general decrease in the number of new positive test results from early April until around May 11th. Afterwards, this trend flattened out and then increased. Over the past two weeks, there has been an average increase of 35 new additional positive cases each day, per day. 

### Positive COVID-19 Tests within the Past Two Weeks
```{r}
recent_df <- 
	sflCases2_df %>% 
	top_n(15, Date)

# Date Range
recentStartEnd_date <- 
	recent_df %>%
	slice(1, n()) %>%
	pull(Date) %>%
	format("%d %B")

# Linear Cases Trend Slope
recentCasesSlope_num <-
	recent_df %>% 
	lm(Positive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)
```

In addition to the overall trend of positive COVID-19 tests shown above, we also show an enhanced view of the last two weeks, highlighting the stark increase in positive cases during this period of time.
```{r cases_over_two_weeks}
nPositiveRecent_gg <- 
	ggplot(data = recent_df) +
	
	theme_bw() +
	theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
	aes(x = Date, y = Positive) + 
	scale_y_continuous(limits = c(0, 1000)) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Number of Positive Test Results within the Past Two Weeks",
		subtitle = paste(
			"Miami Dade County;",
			recentStartEnd_date[1], "to", recentStartEnd_date[2],
			"2020; Linear Change =",
			round(recentCasesSlope_num, 1)
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Count of Positive Cases"
	) +
	
	stat_smooth(method = "loess", colour = "green") +
	geom_point(colour = "green") 

nPositiveRecent_gg
```


## Proportion of Positive Cases within Past Two Weeks
The proportion of positive COVID-19 tests for the past two weeks is:
```{r recent_data}
# Linear Trend Slope
recentSlope_num <- recent_df %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)

# Slope Adjective:
adjective <- case_when(
	abs(recentSlope_num) < 0.1 ~ "slight",
	abs(recentSlope_num) < 0.5 ~ "",
	abs(recentSlope_num) > 0.5 ~ "significant"
)

# Protest Effects: 14 days (incubation period) from George Floyd's death on
#   27 May (see "https://www.cdc.gov/coronavirus/2019-ncov/hcp/clinical-guidance-management-patients.html")
protestEffect_df <- 
	sflCases2_df %>% 
	filter(Date >= "2020-06-10")
protestSlope_num <- 
	protestEffect_df %>% 
	lm(PropPositive ~ Date, data = .) %>% 
	coefficients() %>% 
	pluck(2)
```

```{r recent_proportion}
ggplot(data = recent_df) +
	
	theme_bw() +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Proportion of Positive Test Results within the Past Two Weeks",
		subtitle = paste(
			"Miami Dade County;",
			recentStartEnd_date[1], "to", recentStartEnd_date[2],
			"2020; Linear Change =",
			round(recentSlope_num, 3)
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Proportion (%) of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "black") +
	stat_smooth(method = "lm", colour = "black", se = FALSE, linetype = "dashed")
```

During the 14-day period from `r recentStartEnd_date[1]` to `r recentStartEnd_date[2]`, there was a <b> `r adjective` `r if (recentSlope_num > 0) "increase" else "decrease"` </b> in the proportion of positive COVID-19 cases reported in Miami-Dade County. This increase was even greater during the last week. During the previous week, the proportion of positive cases was increasing close to 1% per day. Given a flat or increasing number of tests performed, the proportion of positive tests obtained is the **key metric** for examining COVID-19 trajectory. **The percentage of positive test results obtained over the past two-weeks has exceeded the 5% positive threshold recommended by the World Health Organization (WHO) for reopening.**


## Proportion of Positive COVID-19 Tests by Day
These are the proportions of positive COVID-19 tests for the since the start of April:
```{r prop_over_time_full}
ggplot(data = sflCases2_df) +
	
	theme_bw() +
	aes(x = Date, y = PropPositive) +
	scale_x_date(
		date_breaks = "1 week",
		date_minor_breaks = "1 day",
		labels = scales::date_format("%d-%b")
	) +
	labs(
		title = "Overall Proportion of Positive Test Results Out of All Tests",
		subtitle = paste(
			"Miami Dade County;",
			startEnd_date[1], "to", startEnd_date[2], "2020"
		),
		caption = "Solid line: LOESS Predictor; Dashed line: Linear Predictor.",
		y = "Proportion (%) of Positive Cases"
	) +
	
	geom_point() +
	stat_smooth(method = "loess", colour = "black") +
	stat_smooth(method = "lm", colour = "black", se = FALSE, linetype = "dashed")
```

Overall, the proportion of positive test results out of all tests **decreased significantly** from over 20% in April to the current proportion of about 12% over the past week. This is function of test availability and the population being tested. However, since the county has begun reopening, the proportion of positive tests is no longer decreasing over time. Given a flat or increasing number of tests performed, an increase in the proportion of positive tests obtained points to evidence of increased COVID-19 community spread in Miami-Dade county.


*******************************************************************************


# Hospitalization Data
```{r hosptitalised}
# Import data
miamidadeHospitalised_df <- read_csv(
	file = "../../data/ESS_MiamiDade_summary_20200620.csv"
)

# Calculate Dates and Save
startEndESS_date <- 
	miamidadeHospitalised_df %>%
	slice(1, n()) %>%
	pull(Date) 


```


## Full Data
```{r}
# Display Data
knitr::kable(miamidadeHospitalised_df) %>% 
	kable_styling() %>%
  scroll_box(height = "500px")
```

</br>

The data above is COVID-19 hospital census data in Miami-Dade county for all dates with available data since April 2. These columns are: the number of patients hospitalized with COVID-19 (**Hospitalised**), the number of patients in intensive care with COVID-19 (**ICU**), the number of patients in intensive care with COVID-19 who are also on ventilators (**Ventilated**), the number of patients positive for COVID-19 that were admitted the previous day (**AdmitPrevDay**), the number of patients recovered from COVID-19 that were discharged the previous day (**DischPrevDay**), and the net admission of patients positive for COVID-19 the previous day (**DeltaAdmit**; a negative value means more recovered patients were discharged than sick patients admitted). *There is a data error for the count ventilated on the 4th of June; we do not know the true value, but we were told verbally by local Department of Health officials that the data point should be approximately 55. We removed this point from the data before estimating the models shown below.*

```{r pivot_hosp_data}
miamidadeHospitalised_df[
	miamidadeHospitalised_df$Date == "2020-06-04",
	"Ventilated"
] <- NA_real_
# There is an error in the original data for June 4th. Rather than deleting the
#   whole row of the raw data, we are skipping this ventilation value

mdCOVID_df <-
	miamidadeHospitalised_df %>%
	pivot_longer(
		Hospitalised:DeltaAdmit,
		names_to = "Type",
		values_to = "Count"
	)
```


## Hospital COVID-19 Census
We now show a census plot of hospitalizations, patients in ICU, and those in ICU on ventilators (shown on a log scale):
```{r graph_hosp_counts, warning = FALSE}
ggplot(
	data = mdCOVID_df %>%
		filter(Type != "DischPrevDay") %>%
		filter(Type != "AdmitPrevDay") %>%
		filter(Type != "DeltaAdmit") 
) +

	theme_bw() +
	theme(legend.position = "bottom") + 
	aes(x = Date, y = Count, group = Type, colour = Type) +
	scale_color_manual(
		values = c(
			"Ventilated" = "#ff0000",
			"ICU" = "#ff7400",
			"Hospitalised" = "#ffc100",
			"AdmitPrevDay" = "black"
		)
	) +
	scale_y_log10() +
	labs(
		title = "Overall Miami-Dade County Hospital COVID-19 Census",
		subtitle = paste(
			"Miami Dade County;",
			format(startEndESS_date[1], "%d %B"),
			"to",
			format(startEndESS_date[2], "%d %B"),
			"2020"
		),
		y = "Counts (Log10 Scale)"
	) +

	geom_point(size = 2) +
	stat_smooth(se = FALSE)
```

Overall, since mid-April, the number of COVID-19 positive patients in Miami-Dade County hospitals has remained constant at around 700 people in hospitals. The COVID-19 ICU census has decreased from its reported peak of 285 on April 17th to around 100 as of June 6th. Overall, the number of COVID-19 patients on ventilators has decreased significantly from its highest point of 192 ventilated on April 22nd to around 60 on June 6th. However, **the observed decreasing trends for ICU census and ventilation census have reversed over the past week and begun to increase for both ICU census and ventilation census, respectively. Additionally, county hospitilizations have begun to increase as well.** Note: data is plotted on a log scale.


*******************************************************************************


# Conclusion

During the past two weeks, the number of COVID-19 tests performed has increased by approximately 35 additional test performed each day, per day. Additionally, there has been a substantial increase in the proportion of positive COVID-19 tests obtained from about 5% in early June, to the current proportion of positive tests of around 12% within the past two weeks. Furthermore, hospitalization census, ICU census, and ventilation census have all begun to increase over approximately the last two weeks. **Taken together, this evidence points to increased community spread of COVID-19 in Miami-Dade County and represents a significant concern for the area.**


*******************************************************************************


# Data Sources
Data sources are as follows:

- **COVID-19 Test Results**: The number of positive and negative COVID-19 test results come from the Florida Department of Health: https://urldefense.proofpoint.com/v2/url?u=https-3A__floridahealthcovid19.gov_&d=DwIGaQ&c=lhMMI368wojMYNABHh1gQQ&r=ZzssdlRrRS7QXnhuTSa63A&m=w5iUGGxeH8o-DvlJLVD9FPmZoJ7DhxHlJ7Aff5hZYVI&s=FRBbMEa8AJCFEWo7Eug19Upwes5IcNte8t5qQPxHmtw&e= .
- **COVID-19 Hospital Census**: The hospitalization data is disseminated by Florida’s Agency for Health Care Administration: https://urldefense.proofpoint.com/v2/url?u=https-3A__ahca.myflorida.com_&d=DwIGaQ&c=lhMMI368wojMYNABHh1gQQ&r=ZzssdlRrRS7QXnhuTSa63A&m=w5iUGGxeH8o-DvlJLVD9FPmZoJ7DhxHlJ7Aff5hZYVI&s=T-_GzcEggFebhnUHIdroUmduqjZCV0jjEkKr9M8WfjY&e= .


